---
output: 
  word_document: 
    fig_caption: yes
    fig_height: 5
    fig_width: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi=400, fig.align = 'center', message = FALSE, warning = FALSE)
# packages used
library(ggplot2)
library(dplyr)
library(ggrepel)
library(reshape)
```


```{r, cache=TRUE}
temp <- read.csv("~/Documents/masters/data/PPOM player probabilities - 2005 to 2017 SG april to april.csv")
temp2 <- temp[temp$Hole.. == 2 | temp$Hole.. == 18, ]
names(temp2) <- c(names(temp2)[-(39:40)], "pwith", "pwithout")
temp2 <- temp2[,c("year", "player", "Hole..", "Round..", "Par", 
                "Score", "RTP.Score", "Level", "pwith")]
df <- tidyr::spread(temp2, Level, pwith)
df <- df[, 1:7]
df$Hole.. <- as.factor(df$Hole..)
names(df)[3] <- "hole"
# group scores greater than 3.
df$score <- ifelse(df$RTP.Score >= 3, 3, df$RTP.Score)
```

Figure 1.  Scoring on Holes 2 and 18 by Year.
```{r Figure 1. Scoring on holes 2 and 18 by year}
# using ggplot
gg <- ggplot(data = df, aes(x = year, y = score)) + 
  geom_jitter(aes(shape = hole, color = hole), size = 1, alpha = 0.2) +
  geom_smooth(aes(group = hole, color = hole), se = FALSE) + 
  scale_x_continuous(breaks = 2005:2017) + ylab("Score to Par") +
  theme_bw() + theme(legend.title = element_blank())
gg
```


```{r}
masters <- read.csv("~/Documents/masters/allyears.csv")
masters$rank <- NULL
# fix some of the ties for first
masters[masters$name == "DiMarco, Chris" & masters$year == 2005, "actualrank"] <- 2
masters[masters$name %in% c("Perry, Kenny", "Campbell, Chad") & masters$year == 2009, "actualrank"] <- 2
masters[masters$name == "Oosthuizen, Louis" & masters$year == 2012, "actualrank"] <- 2
masters[masters$name == "Cabrera, Angel" & masters$year == 2013, "actualrank"] <- 2
masters$winner <- factor(ifelse(masters$actualrank == 1, 1, 0))
# change name to first last for publication purposes
x <- as.character(masters$name)
l <- strsplit(x, ", ")
firstname <- sapply(l, "[", 2)
lastname <- sapply(l, "[", 1)
masters$name <- paste(firstname, lastname)
masters[masters$name == "Dustin Johnson" & masters$year == 2017, "name"] <- "Dustin Johnson (did not play)"
is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}
```


```{r strokes gained data prep}
sg <- masters[masters$year == 2017, c(1, 4:8, 10)]
names(sg) <- c("name", "Approach", "Around", "Tee", "Putting",
               "Total", "OWGR_April")
# remove a few of the players that didn't play very much
sg <- sg[!(sg$name %in% c("Tommy Fleetwood", "Ross Fisher", "Yuta Ikeda")), ]
# code to label univariate outliers
df_melt <- reshape::melt(sg)
df_melt <- df_melt %>%
  group_by(variable) %>%
  mutate(is_outlier = ifelse(is_outlier(value), value, as.numeric(NA)))
df_melt$name[which(is.na(df_melt$is_outlier))] <- as.numeric(NA)
```


Figure 2.  Strokes Gained Data for 2017 Masters.
```{r}
df_sg <- df_melt[df_melt$variable != "OWGR_April", ]
gb <- ggplot(df_sg, aes(x = variable, y = value)) + 
  geom_boxplot(outlier.size = 1.5, outlier.alpha = 0.35, 
               outlier.colour = "red")
gb <- gb + #facet_wrap(~variable, scales = "free", nrow = 2) + 
  theme_bw() + xlab("") + ylab("")
gb <- gb + geom_text_repel(aes(label = name), na.rm = TRUE, 
                           size = 2.5, segment.alpha = 0.5, 
                           segment.size = 0.5)
gb
```


Figure 3.  Official World Golf Ranking of 2017 Masters Participants
```{r, fig.width=4}
## OWGR boxplot
df_owgr <- df_melt[df_melt$variable == "OWGR_April", ]
# need to include DJ
df_owgr$is_outlier <- as.numeric(df_owgr$is_outlier)
go <- ggplot(df_owgr, aes(x = variable, y = value)) + 
  geom_boxplot(outlier.size = 1.5, outlier.alpha = 0.35, 
               outlier.colour = "red")
go <- go + theme_bw() + theme(strip.text.x = element_text(size = 7),
                     axis.title.x = element_blank(),
                     axis.text.x = element_blank(),
                     axis.ticks.x = element_blank())
go <- go + geom_text_repel(aes(label = name), na.rm = TRUE, 
                           size = 2.5, segment.alpha = 0.5, 
                           segment.size = 0.5)
go <- go + ylab("") +xlab("")
go
```


```{r}
df_coef <- read.csv("~/Documents/masters/data/PPOM random player effects.csv")
```

Figure 4a. Random Player Coefficients (histogram)
```{r}
gg <- ggplot(data = df_coef, aes(x = Empirical.Bayes.Estimate)) +
   stat_function(fun = dnorm, 
	args = with(df_coef, c(mean = mean(Empirical.Bayes.Estimate), sd = 0.016)), color = "#FF6666")   
+ geom_histogram(aes(y = ..density..), bins = 30, 
		fill = "white", color = "black", alpha = 0.3) + theme_bw() +
  ylab("") + xlab("")
gg
```


Figure 4b. Random Player Coefficients (boxplot)
```{r, fig.width=4}
df_coef_melt <- reshape::melt(df_coef)
df_coef_est <- df_coef_melt[df_coef_melt$variable ==
                              "Empirical.Bayes.Estimate", ]
# identify univariate outliers
df_coef_est <- df_coef_est %>%
  group_by(variable) %>%
  mutate(is_outlier = ifelse(is_outlier(value), value, as.numeric(NA)))
df_coef_est$name[which(is.na(df_coef_est$is_outlier))] <- as.numeric(NA)

gg <- ggplot(data = df_coef_est, aes(x = variable, y = value)) + 
  geom_boxplot(outlier.size = 1.5, outlier.alpha = 0.35, 
               outlier.colour = "red") + theme_bw() + 
  theme(strip.text.x = element_text(size = 7),
                     axis.title.x = element_blank(),
                     axis.text.x = element_blank(),
                     axis.ticks.x = element_blank()) +
  geom_text_repel(aes(label = name), na.rm = TRUE, 
                           size = 2.5, segment.alpha = 0.5, 
                           segment.size = 0.5) + ylab("") +xlab("")
gg
```


Figure 5.  Probability of Winning vs OWGR
```{r}
df <- masters[, c("name", "year", "owgr.April", "prob", "winner")]
df$is_outlier <- ifelse(df$owgr.April >= 15 | df$winner == 1, 1, NA)
# so the year is displayed
df$name <- paste(df$name, df$year)
df$name[which(is.na(df$is_outlier))] <- as.numeric(NA)
gg <- ggplot(data = df, aes(x = owgr.April, y = prob))
gg <- gg + geom_point(aes(color = winner, shape = winner, 
                          size = winner, alpha = winner)) + 
  theme_bw() + ylab("Predicted Probability of Winning") + 
  xlab("World Golf Ranking in April") + 
  geom_text_repel(aes(label = name), na.rm = TRUE, 
                  size = 2, segment.alpha = 0.5, 
                  segment.size = 0.5) + 
  geom_smooth(se = FALSE, color = "red", alpha = 0.1, size = 0.5,
              lty = 3) + 
  scale_size_manual(values = c(3, 5)) + 
  scale_alpha_manual(values = c(0.35, 1)) + 
  theme(legend.position = "none")
gg
```



Figure 6.  Probability of Winning vs Strokes Gained Total
```{r}
df <- masters[, c("name", "year", "sg.total", "prob", "winner")]
df$is_outlier <- ifelse(df$prob >= 0.3 | df$winner == 1, 1, NA)
# so the year is displayed
df$name <- paste(df$name, df$year)
df$name[which(is.na(df$is_outlier))] <- as.numeric(NA)
gg <- ggplot(data = df, aes(x = sg.total, y = prob))
gg <- gg + geom_point(aes(color = winner, shape = winner, 
                          size = winner, alpha = winner)) + theme_bw() + 
  ylab("Predicted Probability of Winning") + xlab("Strokes Gained Total") + 
  geom_text_repel(aes(label = name), na.rm = TRUE, 
                  size = 2, segment.alpha = 0.5, 
                  segment.size = 0.5) + 
  geom_smooth(se = FALSE, color = "red", alpha = 0.1, size = 0.5,
              lty = 3) + 
  scale_size_manual(values = c(1, 3)) + 
  scale_alpha_manual(values = c(0.5, 1)) + 
  theme(legend.position = "none")
gg
```



Figure 7. Probability of Winning the Masters: Tiger vs Phil
```{r}
tigerphil <- masters[masters$name %in% 
                       c("Tiger Woods", "Phil Mickelson"), ]
gg <- ggplot(data = tigerphil, aes(x = year, y = prob))
gg <- gg + geom_point(aes(shape = name, color = name), size = 4) 
gg <- gg + geom_smooth(aes(group=name, linetype = name, color = name),
                       se = FALSE) + theme_bw()
gg <- gg + scale_x_continuous(breaks = 2005:2017) + 
  ylab("Modeled Probability of Winning") + theme(legend.title = element_blank())
gg
```

